US President Prediction

### 2024-11-04 paper/paper.qmd

```{r}
#| label: fig-harrisposcheck
#| echo: false
#| fig-cap: "Posterior Predictive Check for All Models"
#| message: false
# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots
library(arrow)

@@

pp_check(model1) +  
    theme_classic() +
    theme(legend.position = "bottom")

just_harris_high_quality <- read_parquet(here::here("data/02-analysis_data/Harris.parquet"))

@@

```{r}
#| label: fig-trumpposcheck
#| echo: false
#| fig-cap: "Posterior Predictive Check for All Models"
#| message: false

@@

pp_check(model2) +  
    theme_classic() +
    theme(legend.position = "bottom")

@@

```{r}
#| label: fig-harrisdi
#| echo: false
#| fig-cap: "Model diagonostic for Harris"
#| message: false

@@

plot(model1, "trace")
plot(model1, "rhat")

@@

```{r}
#| label: fig-trumpdi
#| echo: false
#| fig-cap: "Model diagonostic for Trump"
#| message: false

plot(model2, "trace")
plot(model2, "rhat")

@@

## Model Diagnostics {#sec-result-diagnostics}
The model diagnostics indicate a strong fit and reliable predictive performance for both candidates, as detailed in the plots in the model section @sec-model-check. Posterior predictive checks ( @fig-harrisposcheck and @fig-trumpposcheck) show that the predicted values align well with the observed data, suggesting that the models effectively capture key data patterns. Furthermore, the comparison of posterior and prior variables indicates minimal deviation, affirming an appropriate choice of priors. The credible intervals for the predictors (@fig-CIHarris and @fig-CITrump) encompass the observed support rates, demonstrating that the models adequately account for uncertainty in their predictions.

Additional robustness checks (@fig-harrisdi and @fig-trumpdi) confirm the stability of these models. The R-hat values for all parameters are below 1.1, indicating good convergence of the MCMC chains and reliable parameter estimates. This convergence supports the conclusion that the sampling process reached a stable distribution, reinforcing the credibility of our inferences.

>>>>

```{r}
#| label: fig-harrisposcheck_1
#| echo: false
#| fig-cap: "Posterior Predictive Check for Harris model"
#| message: false
# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots
library(arrow)

@@

pp_check(model1) +  
    theme_classic() +
    theme(legend.position = "bottom")
```
```{r}
#| label: fig-harrisposcheck_2
#| echo: false
#| fig-cap: "Comparing the posterior with the prior for Harris model"
#| message: false
just_harris_high_quality <- read_parquet(here::here("data/02-analysis_data/Harris.parquet"))

@@

```{r}
#| label: fig-trumpposcheck_1
#| echo: false
#| fig-cap: "Posterior Predictive Check for Trump model"
#| message: false

@@

pp_check(model2) +  
    theme_classic() +
    theme(legend.position = "bottom")
```

```{r}
#| label: fig-trumpposcheck_2
#| echo: false
#| fig-cap: "Comparing the posterior with the prior for Trump model"
#| message: false

@@

```{r}
#| label: fig-harrisdi_1
#| echo: false
#| fig-cap: "Trace plot for Harris model"
#| message: false

@@

plot(model1, "trace")
```

```{r}
#| label: fig-harrisdi_2
#| echo: false
#| fig-cap: "Rhat plot for Harris model"
#| message: false
plot(model1, "rhat")

@@

```{r}
#| label: fig-trumpdi_1
#| echo: false
#| fig-cap: "Trace plot for Trump model"
#| message: false

# Load necessary libraries
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(patchwork)  # For combining plots
plot(model2, "trace")
```

```{r}
#| label: fig-trumpdi_2
#| echo: false
#| fig-cap: "Rhat plot for Trump model"
#| message: false
plot(model2, "rhat")

@@

## Model Diagnostics {#sec-result-diagnostics}
The model diagnostics indicate a strong fit and reliable predictive performance for both candidates, as detailed in the plots in the model section @sec-model-check. Posterior predictive checks ( @fig-harrisposcheck_1, @fig-harrisposcheck_2, @fig-trumpposcheck_1, and @fig-trumpposcheck_2) show that the predicted values align well with the observed data, suggesting that the models effectively capture key data patterns. Furthermore, the comparison of posterior and prior variables indicates minimal deviation, affirming an appropriate choice of priors. The credible intervals for the predictors (@fig-CIHarris and @fig-CITrump) encompass the observed support rates, demonstrating that the models adequately account for uncertainty in their predictions.

Additional robustness checks (@fig-harrisdi_1, @fig-harrisdi_1, @fig-trumpdi_1, and @fig-trumpdi_2) confirm the stability of these models. The R-hat values for all parameters are below 1.1, indicating good convergence of the MCMC chains and reliable parameter estimates. This convergence supports the conclusion that the sampling process reached a stable distribution, reinforcing the credibility of our inferences.
